### 2020-01-22 articletwo-sensitivityanalysis.Rmd
init

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

### Plot functions

<details>
  <summary>Open/Close</summary>

```{r Plot functions, code=readLines("articletwo-ggplot.R")}
source("articletwo-ggplot.R")
```

</details>

# Import previous session

```{r Import previous session}
pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
names.dset <- getdescriptions()
```

## Variables

```{r my variables}
var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
var.CL.min <- c("BL_AGE", "SEX")
var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
var.CL.opt <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB", "BP_TREAT")
permutations <- 99
```

# Sensitivity analysis

```{r sensitivity analysis bp treat}
adonis.species.bptreat <- calculate.betadiversity(pseq = pseq.species,
                                          matrix = bray.dist.m.species,
                                          vars = list("max" = var.CL.opt,
                                                      "min" = var.CL.min))

diversity.bptreat <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL.opt, "min" = var.CL.min),
                                 betadiversity = adonis.species.bptreat,
                                 names.dset = names.dset)

g.diversity.bptreat <- plot.diversities(diversity.bptreat)
ggsave(file = "cache/sensitivity-alphabeta-bptreat.png", plot = g.diversity.bptreat, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-bptreat.png"/>

```{r sensitivity analysis only healthy}
pseq.species.healthy <- subset_samples(pseq.species,
                                       PREVAL_DIAB == 0 &
                                       PREVAL_CHD == 0 &
                                       PREVAL_CR_ANYCANC == 0)

bray.dist.m.species.healthy <- bray.dist.m.species[pseq.species.healthy %>% meta %>% rownames,
                                                   pseq.species.healthy %>% meta %>% rownames]

adonis.species.healthy <- calculate.betadiversity(pseq = pseq.species.healthy,
                                          matrix = bray.dist.m.species.healthy,
                                          vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                      "min" = var.CL.min))

diversity.healthy <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                             "min" = var.CL.min),
                                 betadiversity = adonis.species.healthy,
                                 names.dset = names.dset)

g.diversity.healthy <- plot.diversities(diversity.healthy)
ggsave(file = "cache/sensitivity-alphabeta-healthy.png", plot = g.diversity.healthy, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-healthy.png"/>

## BMI

Normal weigth

```{r sensitivity analysis normal weight}
pseq.species.normalweight <- subset_samples(pseq.species, BMI > 18.5 & BMI <= 25)

bray.dist.m.species.normalweight <-
    bray.dist.m.species[pseq.species.normalweight %>% meta %>% rownames,
                        pseq.species.normalweight %>% meta %>% rownames]

adonis.species.normalweight <- calculate.betadiversity(pseq = pseq.species.normalweight,
                                          matrix = bray.dist.m.species.normalweight,
                                          vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                      "min" = var.CL.min))

diversity.normalweight <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                             "min" = var.CL.min),
                                 betadiversity = adonis.species.normalweight,
                                 names.dset = names.dset)

g.diversity.normalweight <- plot.diversities(diversity.normalweight)
ggsave(file = "cache/sensitivity-alphabeta-normalweight.png",
       plot = g.diversity.normalweight, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-normalweight.png"/>

```{r sensitivity analysis obese}
pseq.species.obese <- subset_samples(pseq.species, BMI > 30 & BMI <= 40)

bray.dist.m.species.obese <-
    bray.dist.m.species[pseq.species.obese %>% meta %>% rownames,
                        pseq.species.obese %>% meta %>% rownames]

adonis.species.obese <- calculate.betadiversity(pseq = pseq.species.obese,
                                          matrix = bray.dist.m.species.obese,
                                          vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                                      "min" = var.CL.min))

diversity.obese <- diversities(pseq = pseq.species,
                                 vars = list("max" = var.CL[var.CL != "PREVAL_DIAB"],
                                             "min" = var.CL.min),
                                 betadiversity = adonis.species.obese,
                                 names.dset = names.dset)

g.diversity.obese <- plot.diversities(diversity.obese)
ggsave(file = "cache/sensitivity-alphabeta-obese.png", plot = g.diversity.obese, height = 6, width = 9)
```

<img src="cache/sensitivity-alphabeta-obese.png"/>

```{r save session}
save.image(file = "session/session-sensitivity.Rdata")
```

### 2020-01-22 articletwo.Rmd
init

---
title: "Hypertension and microbiome"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)

dir.create("cache/", showWarnings = FALSE)
dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)

now <- format(Sys.time(), '%Y%m%d-%H%M%S')
```

```{r Command line arguments, include = FALSE, eval = FALSE}
file <- paste0("session/", sort(list.files("session"), decreasing = TRUE)[1])
message("Loading variables from file ", file)
load(file)
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(egg)
library(gridExtra)
library(png)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
library(DESeq2)
library(ggeffects)
library(pander)
library(hexbin)
```

Session info

```{r Session info}
pander(sessionInfo(), compact = TRUE)
```

# Sources

<details><summary>Functions</summary>

```{r Functions}
sourcefiles <- c("articletwo-officer.R",
                 "articletwo-rrbiome.R",
                 "articletwo-ggplot.R")
```

```{r import files, echo = FALSE}
for (f in sourcefiles) {
    source(f)
}
```

```{r embbed files, echo = FALSE}
xfun::embed_files(c("rrnmr.Rmd", sourcefiles))
```

</details>

# Loading data

Loading descriptions for clinical data

```{r variables, warning = FALSE}
names.dset <- getdescriptions()
```

```{r temp}
temp <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
temp %>% meta %>% nrow
```

Loading phyloseq object

```{r data}
pseq.species <- import_filter_data("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
pseq.genus <- import_filter_data("data/phfinrisk_genus_all_drop50k_2018-11-16.RDs")
pseq.genus.coretaxa <- coretaxa(pseq.genus, detection = 0.1/100, prevalence = 1/100)
```

Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
genus `r pseq.genus %>% sample_sums %>% mean`. Core has length 
`r length(pseq.genus.coretaxa)`.

At species level meta has dimensions (`r dim(meta(pseq.species))`) and
there ntaxa is `r ntaxa(pseq.species)`. At genus level meta has
dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
`r ntaxa(pseq.genus)`.

# Variables

```{r my variables}
var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
var.CL.min <- c("BL_AGE", "SEX")
var.CL <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
```

# Heavy models

### Bray curtis distance matrix

```{r matrix calculation}
if (!file.exists("rds/bray.dist.m.species.rds")) {
    bray.dist.m.species <- calculate.beta.matrix(pseq.species)
    saveRDS(bray.dist.m.species, file = "rds/bray.dist.m.species.rds")
} else {
    bray.dist.m.species  <- readRDS("rds/bray.dist.m.species.rds")
}
```

### Beta diversity

 ```{r adonis calculation}
 if (!file.exists("rds/adonis.species.rds")) {
     adonis.species <- calculate.betadiversity(pseq = pseq.species,
                                               matrix = bray.dist.m.species,
                                               vars = list("max" = var.CL,
                                                            "min" = var.CL.min))
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
 } else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
 }
```

### PCoA

```{r pcoa calculate}
if (!file.exists("rds/pcoa.ordinate.rds")) {
    pcoa.abundances <- microbiome::transform(pseq.species, 'compositional')
    pcoa.ordinate <- ordinate(pcoa.abundances, method="PCoA", distance="bray")
    saveRDS(pcoa.ordinate, file = "rds/pcoa.ordinate.rds")
} else {
    pcoa.ordinate <- readRDS("rds/pcoa.ordinate.rds")
}
```

### DeSeq2

```{r deseq2}
if (!file.exists("rds/dds.rds") || !file.exists("rds/dds3.rds") || !file.exists("rds/ddsNa.rds")) {
    pseq.genus.core.deseq <-  prune_taxa(pseq.genus.coretaxa, pseq.genus)

    if (!file.exists("rds/dds.rds")) {
        dds <- lapply(c2l(var.BP), function(x, pseq) {
            dds.data <- phyloseq_to_deseq2(pseq, deseq.formula(x, var.CL))
            DESeq(dds.data,
                  test="Wald",
                  fitType="parametric",
                  parallel = TRUE,
                  BPPARAM=MulticoreParam(16))
        }, pseq = pseq.genus.core.deseq)
        saveRDS(dds, file = "rds/dds.rds")
    }
    if (!file.exists("rds/dds3.rds")) {
        dds3 <- lapply(deseq.list(var.CL, var.CL.min), function(x, pseq) {
            dds.data <- phyloseq_to_deseq2(pseq, deseq.formula(x, "HYPERTENSION"))
            DESeq(dds.data,
                  test="Wald",
                  fitType="parametric",
                  parallel = TRUE,
                  BPPARAM=MulticoreParam(16))
        }, pseq = pseq.genus.core.deseq)
        saveRDS(dds3, file = "rds/dds3.rds")
    }
    if (!file.exists("rds/ddsNa.rds")) {
        ddsNa <- mclapply(c2l(var.BP), function(x) {
            myDESeq(pseq = pseq.species,
                    vars = c(x, var.CL),
                    coreterm = "^Lactobacillus.*Bacteria")
        }, mc.cores = length(var.BP))
        saveRDS(ddsNa, file = "rds/ddsNa.rds")        
    }
} else {
    dds <- readRDS(file = "rds/dds.rds")
    dds3 <- readRDS(file = "rds/dds3.rds")
    ddsNa <- readRDS(file = "rds/ddsNa.rds")
}
```

# Results

## Characteristics

```{r Characteristics}
characteristicsTable(meta(pseq.species)) %>% kable
```

<details><summary>Salt subset</summary>

```{r Characteristics subset, echo = FALSE}
pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE) %>%
    meta %>%
    characteristicsTable %>%
    kable
```

</details>

## Microbial diversity

```{r alphabeta definitions}
diversity <- diversities(pseq = pseq.species,
                           vars = list("max" = var.CL, "min" = var.CL.min),
                           betadiversity = adonis.species,
                           names.dset = names.dset)
```

Results in minimum model of alpha diversity

```{r talbe alpha div min}
diversities.tidy(diversity) %>%
    kable
```

```{r save grob}
g.diversity <- plot.diversities(diversity)
ggsave(file = "cache/alpha-beta.png", plot = g.diversity, height = 6, width = 9)
```

<img src="cache/alpha-beta.png" />

## Principal coordinate analysis

```{r pcoa plot}
pcoa.plot <- pcoaplot(pcoa.ordinate)
ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.0, width=6, units = "in", dpi = 300)
```

<img src="cache/pcoa-species.png" />

## Associations for genera

```{r deseq continuous}
dset.deseq.signf <- deseqresults(dds, names.dset)
g.deseq <- deseqheatmap(dset.deseq.signf)
```

Number of significant associations `r sum(pull(dset.deseq.signf, qval) < 0.05)` for 
`r pull(dset.deseq.signf, Feature)  %>% unique %>% length` features.

### Number of associations per covariate

```{r results compare deseq2}
map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)) %>%
    map_df(~as.data.frame(.x) %>% dplyr::rename(qval = .x), .id = "included") %>%
    kable
```

## Associations for Lactobacillus species

```{r lactobacillus species results}
dset.deseq.signf.na <- deseqresults(ddsNa, names.dset)
g.deseq.na <- deseqheatmap(dset.deseq.signf.na)
```

Number of significant associations `r sum(pull(dset.deseq.signf.na, qval) < 0.05)` for 
`r dset.deseq.signf.na %>% nrow` features. Number of
`r dset.deseq.signf.na %>% filter(log2FoldChange > 0) %>% nrow`
association were positive.

```{r deseq dual plot}
g.deseqgenus <- gtable_frame(ggplotGrob(g.deseq + theme(legend.position = "none")),
                  height = unit(length(unique((dset.deseq.signf$Feature))), 'mm'),
                  width = unit(25, 'mm'))

g.deseqspecies <- gtable_frame(ggplotGrob(g.deseq.na + theme(legend.position = c(0.7, 1.45))),
                  height = unit(length(unique((dds.results$Feature))), 'mm'),
                                width = unit(25, 'mm'))

g.deseq <- arrangeGrob(g.deseqgenus, g.deseqspecies, ncol = 2)

ggsave(file = "cache/deseq-full.png", plot = g.deseq, width = 14, height = 25, dpi = 300, unit = "cm")
```

<img src="cache/deseq-full.png" />

## Salt subset

## Association between genera and dU-NA

```{r lacto deseq calculation}
dds.salt <- myDESeq(pseq = pseq.genus,
                    coretaxa = pseq.genus.coretaxa,
                    vars = c("NA.", var.CL),
                    saltsubset = TRUE)
```

Lactobacillus prevalence in subsample

```{r lactobacillus prevalence}
pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
pseq_prevalence(pseq.genus.salt)
```

```{r significant results for salt subset}
dds.salt %>%
    results(., name = "NA.", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
    filter(row == "Lactobacillus (Bacteria)") %>%
    kable
```

Graphical representation for findings

```{r scatter plot from deseq}
g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
ggsave(file = "cache/gsalt.png", plot = g.salt, height = 3, width = 3, dpi = 300)
```

<img src="cache/gsalt.png" />

## Association between Lactobacilli and dU-NA

```{r species salt}
dds.salt.species <- myDESeq(pseq = pseq.species,
                        vars = c("NA.", var.CL),
                        coreterm = "^Lactobacillus.*Bacteria",
                        saltsubset = TRUE)
```

Significant results after FDR correction

```{r species salt results}
dds.salt.species %>%
    results(., name = "NA.", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
                  row = renametaxa(row)) %>%
    filter(qval < 0.05) %>%
    kable
```

# Supplements

Listing our "core" genera

```{r write supplement}
core.names.fourcols <- pseq.genus.coretaxa %>%
    renametaxa %>%
    sort %>%
    list.partition %>%
    mycbind

pseq.genus.coretaxa %>% length

core.names.flextable <- flextable(core.names.fourcols) %>%
    flextable::fontsize(size = 10, part = "header") %>%
    flextable::fontsize(size = 10, part = "body") %>%
    flextable::width(j=1:4, width = 1.8) %>%
    flextable::align(align = "left", part = "header") %>%
    flextable::align(align = "left")
core.names.flextable
```

Listing results for alpha diversity

```{r supplement table for alpha diversity}
alphadiversity.table <- diversity %>% map_df(~as.data.frame(.x), .id = "model") %>%
    mutate(mean_ci = sprintf("%.2f (%.2f - %.2f)", alpha.effect, alpha.low, alpha.high),
           pubp = pub.p(alpha.p)) %>%
    myspread(list = c2l("mean_ci", "pubp"), term = "Name", key = "model") %>%
    arrange(match(Name, c("Systolic BP", "Diastolic BP", "Mean arterial pressure", "Pulse pressure", "Hypertension"))) %>%
        select(Name, min_mean_ci, min_pubp, max_mean_ci, max_pubp)

typology.tbls2 <- data.frame(
    col_keys = colnames(alphadiversity.table),
    what = c("", rep("Age- and sex adjusted model", 2), rep("Multivaraible adjusted model", 2)),
    measure = c("", rep(c("beta (95%-CI)", "p"), 2)),
    stringsAsFactors = FALSE)

alphadiversity.flextable <-
    typologyformatter(data = alphadiversity.table, font = 12, typology = typology.tbls2) %>%
    flextable::width(j=1, width = 1.9) %>%
    flextable::width(j=c(2:2,4:4), width = 1.8) %>%
    flextable::width(j=c(3:3,5:5), width = 0.7)
alphadiversity.flextable 
```

```{r supplement table for beta diversity}
betadiversity.table <- diversity %>% map_df(~as.data.frame(.x), .id = "model") %>%
    mutate(rsquared = sprintf("%.3f%%", beta.R2*100),
           pubp = pub.p(beta.p)) %>%
    myspread(list = c2l("rsquared", "pubp"), term = "Name", key = "model") %>%
    arrange(match(Name, c("Systolic BP", "Diastolic BP", "Mean arterial pressure", "Pulse pressure", "Hypertension"))) %>%
    select(Name, min_rsquared, min_pubp, max_rsquared, max_pubp)

typology.tbls3 <- data.frame(
    col_keys = colnames(betadiversity.table),
    what = c("", rep("Age- and sex adjusted model", 2), rep("Multivaraible adjusted model", 2)),
    measure = c("", rep(c("R2", "p"), 2)),
    stringsAsFactors = FALSE)

betadiversity.flextable <-
    typologyformatter(data = betadiversity.table, font = 12, typology = typology.tbls3) %>%
    flextable::width(j=1, width = 1.9) %>%
    flextable::width(j=c(2:5), width = 0.8)

betadiversity.flextable
```

```{r supplement table deseq}
typology.tbls4 <- data.frame(
    col_keys = colnames(deseq.table),
    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
    stringsAsFactors = FALSE)

deseq.table <- dset.deseq.signf %>%
    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
           p.value = pub.p(qval)) %>%
    myspread %>%
    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("Pulse pressure"),
           starts_with("MAP"), starts_with("Hypertension"))

typologyformatter(data = deseq.table, font = 9, typology = typology.tbls4) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55)
#  %>%   save_as_html("cache/genus.html")
```

```{r supplement table deseq}
deseq.table.na <- dset.deseq.signf.na %>%
    mutate(lfc_se = sprintf("%.2f±%.2f", log2FoldChange, lfcSE),
           p.value = pub.p(qval)) %>%
    myspread %>%
    select(Feature, starts_with("Systolic"), starts_with("Diastolic"), starts_with("Pulse pressure"),
           starts_with("MAP"), starts_with("Hypertension"))

typologyformatter(data = deseq.table.na, font = 9, typology = typology.tbls4) %>%
    flextable::width(j = 1, width = 1.2) %>%
    flextable::width(j = seq(2, 11, 2), width = 0.75) %>%
    flextable::width(j = seq(3, 11, 2), width = 0.55) 
#  %>%   save_as_html("cache/species.html")
```

```{r save session}
save.image(file = paste0("session/session-", now, ".Rdata"))
```

### 2020-01-22 articletwo-ggplot.R

@@ -1,0 +1,321 @@
+plot.alpha <- function(diversity,
+                       alphabreaks = seq(-1, 0, 0.2),
+                       alphalabels = c("(-1.0, -0.8]", "(-0.8, -0.6]", "(-0.6, -0.4]", "(-0.4, -0.2]", "(-0.2, -0.0]")) {
+    diversity <- diversity %>%
+        mutate(estimate = ifelse(response == "HYPERTENSION", log(alpha.effect), alpha.effect),
+               estimate_fac = cut(estimate,
+                                  breaks = alphabreaks,
+                                  labels = alphalabels,
+                                  include.lowest = TRUE),
+               Qstar = ifelse(alpha.p < 0.05, '*', ' '))
+
+    plot.alpha.diversity <- ggplot(diversity, aes(x = reorder(Name, -deseqplotorder(Name)),
+                                                  y = 1,
+                                                  fill = estimate_fac)) +
+        geom_bar(stat="identity", color="black") +
+        coord_flip() +
+        scale_fill_brewer(palette="Blues",
+                          name = 'Regression coefficient \nin linear model \nfor Shannon index',
+                          drop=FALSE,
+                          direction = -1) +
+        theme_classic(20) +
+        geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=14) +
+        scale_shape_manual(name="",
+                           values=c('*'='*', ' '=' '),
+                           labels=c("*"="significant at\nFDR 0.05", ' '=' '),
+                           breaks=c("*", ' ')) +
+        xlab("") +
+        ylab("") +
+        theme(legend.position="right",
+              legend.title=element_text(size = 14),
+              legend.direction="vertical",
+              legend.text=element_text(size = floor(0.75*20)),
+              legend.box.margin=margin(t = 0, unit = "cm"),
+              legend.margin=margin(0,0,0,0,"pt"),
+              legend.justification="left",
+              axis.line.y = element_line(linetype="blank"),
+              axis.line.x = element_line(linetype="blank"),
+              legend.title.align = 0,
+              legend.text.align = 0,
+              axis.text.y = element_text(color="black"),
+              axis.text.x = element_text(color="black"))+
[…]

### 2020-01-22 articletwo-officer.R

@@ -1,0 +1,119 @@
+getfactorvariables <- function(df, vars) {
+    classes <- lapply(c2l(vars), function(x) class(df[[x]]))
+    classes[classes == "factor"] %>% names
+}
+
+characteristics.names <- function(onlyvars = FALSE) {
+    if (onlyvars == TRUE)
+        return(c("BL_AGE", "SEX", "BMI", "oSYSTM", "oDIASM",
+                 "oPULSEPRESSURE", "oMAP", "HYPERTENSION",  
+                 "CURR_SMOKE", "PREVAL_DIAB",  "Q57X", "BL_USE_RX_C03",
+                 "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"))
+    list("BL_AGE" = "Age, y (SD)",
+         "SEX" = "Female, N (%)",
+         "BMI" = "BMI, kg/m² (SD)",
+         "oSYSTM" = "Systolic BP, mmHg (SD)",
+         "oDIASM" = "Diastolic BP, mmHg (SD)",
+         "oPULSEPRESSURE" = "Pulse pressure, mmHg (SD)",
+         "oMAP" = "Mean arterial pressure, mmHg (SD)",
+         "HYPERTENSION" = "Hypertension, N (%)",
+         "CURR_SMOKE" = "Current smoker, N (%)",
+         "PREVAL_DIAB" = "Diabetes mellitus, N (%)",
+         "Q57X" = "Exercise, N (%)",
+         "ANYDRUG" = "Antihypertensive medication, N (%)",
+         "BL_USE_RX_C03" = "  Diuretics, N (%)",
+         "BL_USE_RX_C07" = "  Beta blockers, N (%)",
+         "BL_USE_RX_C08" = "  Calcium channel blockers, N (%)",
+         "BL_USE_RX_C09" = "  RAS blockers, N (%)",
+         "1"  =  "  Light",
+         "2"  =  "  Moderate",
+         "3"  =  "  Heavy",
+         "4"  = "  Competitive")
+}
+
+characteristicsTable <- function(dset, strata) {
+    nostrata <- missing(strata)
+    characteristics <- tableone::CreateTableOne(
+                                     strata = strata,
+                                     data = dset,
+                                     vars = characteristics.names(TRUE),
+                                     factorVars = getfactorvariables(dset,
+                                                                     characteristics.names(TRUE)),
+                                     test = !missing(strata))
+    print(characteristics,
+                      exact = "stage",
[…]

### 2020-01-22 articletwo-rrbiome.R

@@ -1,0 +1,420 @@
+# Helper functions for articletwo
+
+replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
+
+renametaxa <- function (names) {
+    suffix <- ifelse(grepl("BacteriaPlasmid", names), "*", "")
+    base <- ifelse(grepl("_", names),
+           gsub('(.).*_(.*) \\(+(.+)\\)', '\\1. \\2', names),
+           gsub('(.*) \\(+(.+)\\)', '\\1', names))
+    paste0(base, suffix)
+}
+
+mycbind <- function(list) {
+    max.length <- c(list) %>% lapply(length) %>% unlist %>% max
+    lapply(list, function(x) {
+        length(x) <- max.length
+        x}) %>%
+        do.call(cbind, .)
+}
+
+list.partition <- function(list, parts = 3) {
+    partition <- floor(parts*seq(0, length(list)-1)/length(list))
+    split(list, partition)
+}
+
+rename.genus  <- function (genus, markword = "Plasmid", mark = "*") {
+    name <- gsub('(.*) \\(.*\\)', '\\1', genus)
+    star <- ifelse(grepl(markword, genus), mark, "")
+    paste0(gsub("_", " ", name), star)
+}
+
+myscale <- function(x){
+  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
+}
+
+filter.phenotype.data <- function(pseq,
+                                  included = c("MEN",
+                                               "BL_AGE",
+                                               "SYSTM",
+                                               "DIASM",
+                                               "BMI",
+                                               "CURR_SMOKE",
+                                               "PREVAL_DIAB",
+                                               "Q57X",
+                                               "BL_USE_RX_C09",
+                                               "BL_USE_RX_C03",
+                                               "BL_USE_RX_C07",
+                                               "BL_USE_RX_C08",
+                                               "NA.",
+                                               "BP_TREAT",
+                                               "PREVAL_CHD",
+                                               "PREVAL_CR_ANYCANC"),
+                                  allowna = c("NA.",
[…]

### 2020-01-30 articletwo.Rmd
Most recent

@@
 names.dset <- getdescriptions()
-```
-
-```{r temp}
-temp <- readRDS("data/phfinrisk_species_all_drop50k_2018-12-21.RDs")
-temp %>% meta %>% nrow
 ```
@@
 characteristicsTable(meta(pseq.species)) %>% kable
-```
+``
 
@@
 g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
-ggsave(file = "cache/gsalt.png", plot = g.salt, height = 3, width = 3, dpi = 300)
+ggsave(file = "cache/gsalt.png", plot = g.salt, height = 5, width = 5, dpi = 300)
 ```
@@
 ```{r write supplement}
-core.names.fourcols <- pseq.genus.coretaxa %>%
+core.names.incols <- pseq.genus.coretaxa %>%
     renametaxa %>%
@@
     list.partition %>%
-    mycbind
+    mycbind %>%
+    as.data.frame
 
-pseq.genus.coretaxa %>% length
-
-core.names.flextable <- flextable(core.names.fourcols) %>%
+core.names.flextable <- flextable(core.names.incols) %>%
     flextable::fontsize(size = 10, part = "header") %>%
     flextable::fontsize(size = 10, part = "body") %>%
-    flextable::width(j=1:4, width = 1.8) %>%
+    flextable::width(j=1:3, width = 1.8) %>%
     flextable::align(align = "left", part = "header") %>%
     flextable::align(align = "left")
-core.names.flextable
+
+core.names.flextable %>% save_as_html("cache/s1.html")
 ```
@@
 
-```{r supplement table deseq}
+```{r supplement table deseq na}
 deseq.table.na <- dset.deseq.signf.na %>%

### 2020-01-30 maaslinruns.Rmd
Most recent

---
title: "Maaslin runs"
author: "Joonatan Palmu"
date: "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, xecho = TRUE, message = FALSE, results='asis',
                      cache=FALSE, warning=FALSE)
knitr::opts_chunk$set(cache.path = 'cache/', output.dir="cache/",
                      file.path = 'cache/', fig.path = 'cache/')

options(max.print=100)

dir.create("rds/", showWarnings = FALSE)
dir.create("session/", showWarnings = FALSE)
```

# Command line arguments

Calculations ran at

```{r Command line arguments}
now <- format(Sys.time(), '%Y%m%d-%H%M%S')

if (exists("args")) {
    if ("time" %in% names(args)) now <- args$time

    if("clean" %in% names(args)) {
        unlink("rds", recursive=TRUE)
        unlink("cache", recursive=TRUE)
    }
    if ("tags" %in% names(args)) 
        rtags('~/phd/research/articletwo/', recursive = FALSE, pattern = '\\.[RrSs](rw|md)?$',
              ofile = '~/phd/research/articletwo/TAGS', verbose = TRUE, append = FALSE)
    if ("loadlast" %in% names(args) )
        load(paste0("session/", sort(list.files("session"), decreasing = TRUE)[1]))
    if ("subset" %in% names(args))
        subset.run <- as.numeric(args$subset)
}

#packrat::opts$ignored.packages
```

# Importing libraries:

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
```

### RR biome functions

<details>
  <summary>Open/Close</summary>

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```

</details>

```{r data}
pseq.genus <- import_filter_data("data/phfinrisk_centrifuge_genus_d50k_2019-06-02.rds")
pseq.genus.core <- microbiome::transform(pseq.genus, "compositional") %>%
    core(detection = 0.0001, prevalence = 0.01)
```

## Variables

```{r my variables}
var.BP <- c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")
var.CL.min <- c("BL_AGE", "SEX")
var.CL.add <- c("SEX", "BMI", "CURR_SMOKE", "Q57X", "PREVAL_DIAB",
            "BL_USE_RX_C03","BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09")
```

```{r var combinations}
var.CL.list <- seq(length(var.CL.add)) %>% as.list
names(var.CL.list) <- var.CL.add
var.CL.comb <- lapply(var.CL.list, function(x) unique(c(var.CL.min, var.CL.add[1:x])))
var.CL.comb3rd <- lapply(var.CL.list, function(x) unique(c(var.CL.min, var.CL.add[x])))
```

```{r maaslin runs, results = 'hide', eval  = FALSE} 
ret <- mclapply(var.CL.comb, function(x)
    maaslinwrapper(pseq = pseq.genus.core,
                   looped = var.BP,
                   forced = x,
                   taxa = taxa(pseq.genus.core),
                   tempstr = sprintf("%%s/maaslinruns/combinations/single_%s", paste0(x, collapse = "_"))),
    mc.cores = length(var.CL.comb))
```

```{r adding third, results = 'hide', eval = FALSE}
ret3rd <- mclapply(var.CL.comb3rd, function(x)
    maaslinwrapper(pseq = pseq.genus.core,
                   looped = var.BP,
                   forced = x,
                   taxa = taxa(pseq.genus.core),
                   tempstr = sprintf("%%s/maaslinruns/combinations/all3rd_%s", paste0(x, collapse = "_"))),
    mc.cores = length(var.CL.comb3rd))

```

```{r read combinations}
ret <- lapply(var.CL.comb, function(x) {
    file <- sprintf("/homes/jpalmu/maaslinruns/combinations/single_%s/maaslin/maaslin.txt",
                    paste0(x, collapse = "_"))
    read.table(file, header=TRUE, sep='\t') %>% as.data.frame
    })
```

```{r combine results}
ret.comb <- map_df(ret, ~as.data.frame(.x), .id="id")

ret.comb %>%
    dplyr::filter(Q.value < 0.05) %>%
    group_by(id) %>%
    summarize(n = n(), model = paste0(var.CL.comb[[id[1]]], collapse = "+")) %>%
    arrange(model) %>%
    kable
```

```{r read combinations 3rd}
ret3rd <- lapply(var.CL.comb3rd, function(x) {
    file <- sprintf("/homes/jpalmu/maaslinruns/combinations/all3rd_%s/maaslin/maaslin.txt",
                    paste0(x, collapse = "_"))
    read.table(file, header=TRUE, sep='\t') %>% as.data.frame
    })
```

```{r combine results 3rd}
ret3rd.comb <- map_df(ret3rd, ~as.data.frame(.x), .id="id")

ret3rd.comb %>%
    dplyr::filter(Q.value < 0.05) %>%
    group_by(id) %>%
    summarize(n = n(), model = paste0(var.CL.comb3rd[[id[1]]], collapse = "+")) %>%
    arrange(n) %>%
    kable
```

### 2020-01-30 articletwo-ggplot.R

@@ -229,7 +229,8 @@ saltplot <- function(dset, ymax = 2) {
         theme_classic() +
         theme(plot.title = element_blank(),
               legend.title = element_blank(),
-              legend.position = c(0.9,0.95))
+              legend.position = c(0.9,0.95),
+              axis.text = element_text(size = 8))
 }
 
 myoutlier <- function(list) {
@@ -267,7 +268,8 @@ saltboxplot <- function(pseq, dds) {
         ylab("24-hour urinary sodium [mmol]") +
         theme_classic() +
         theme(strip.background = element_blank(),
-              strip.placement = "outside")
+              strip.placement = "outside",
+              axis.title = element_text(size = 9))
 }
 
 pcoaplot <- function(pcoa) {

### 2020-01-30 articletwo-officer.R

@@ -7,7 +7,7 @@ characteristics.names <- function(onlyvars = FALSE) {
     if (onlyvars == TRUE)
         return(c("BL_AGE", "SEX", "BMI", "oSYSTM", "oDIASM",
                  "oPULSEPRESSURE", "oMAP", "HYPERTENSION",  
-                 "CURR_SMOKE", "PREVAL_DIAB",  "Q57X", "BL_USE_RX_C03",
+                 "CURR_SMOKE", "PREVAL_DIAB",  "Q57X", "ANYDRUG", "BL_USE_RX_C03",
                  "BL_USE_RX_C07", "BL_USE_RX_C08", "BL_USE_RX_C09"))
     list("BL_AGE" = "Age, y (SD)",
          "SEX" = "Female, N (%)",

### 2020-02-01 articletwo.Rmd
For reviewer

@@
 date: "`r format(Sys.time(), '%d.%m.%Y')`"
-output: html_document
+output: 
+  html_document:
+    number_sections: true
 ---
 
-```{r setup, include=FALSE}
+```{r setup, echo=FALSE}
 knitr::opts_chunk$set(include = TRUE, echo = TRUE, message = FALSE, results='asis',
@@
 
-options(max.print=100)
+options(knitr.kable.NA = "")
 
@@
 
-```{r Command line arguments, include = FALSE, eval = FALSE}
+```{r Command line arguments, echo = FALSE, eval = FALSE}
 file <- paste0("session/", sort(list.files("session"), decreasing = TRUE)[1])
@@
 # Importing libraries:
+
+<details><summary>Libraries</summary>
 
@@
 
-Session info
+</details>
+
+# Session info
+
+<details><summary>Session</summary>
 
@@
 ```
+
+</details>
 
@@
 
-Species has average number of reads `r pseq.species %>% sample_sums %>% mean` and
-genus `r pseq.genus %>% sample_sums %>% mean`. Core has length 
-`r length(pseq.genus.coretaxa)`.
-
 At species level meta has dimensions (`r dim(meta(pseq.species))`) and
@@
 dimensions (`r dim(meta(pseq.genus))`) and there ntaxa is 
-`r ntaxa(pseq.genus)`.
+`r ntaxa(pseq.genus)`. Core size is `r length(pseq.genus.coretaxa)`.
 
@@
 characteristicsTable(meta(pseq.species)) %>% kable
-``
+```
 
@@
 
-Results in minimum model of alpha diversity
-
-```{r talbe alpha div min}
+```{r talbe alpha div min, echo = FALSE}
 diversities.tidy(diversity) %>%
@@
 
-```{r save grob}
+Diversity plot
+
+```{r save grob, echo = FALSE}
 g.diversity <- plot.diversities(diversity)
@@
 
-```{r pcoa plot}
+```{r pcoa plot, echo = FALSE}
 pcoa.plot <- pcoaplot(pcoa.ordinate)
@@
 
-Number of significant associations `r sum(pull(dset.deseq.signf, qval) < 0.05)` for 
-`r pull(dset.deseq.signf, Feature)  %>% unique %>% length` features.
+Significant associations found between blood pressure indices and N:o `r nrow(dset.deseq.signf)`
+genera and N:o `r dset.deseq.signf %>% pull(Feature) %>% unique %>% length` features. N:o 
+`r dset.deseq.signf %>% filter(log2FoldChange > 0) %>% nrow` positive associations found.
 
-### Number of associations per covariate
+### Associations for hypertension looping covariates
+
+Base model had 
+`r sum(results(dds3$BL_AGE, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`
+significant associations and inclusion of BMI dropped it to
+`r sum(results(dds3$BMI, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)`.
 
 ```{r results compare deseq2}
-map(dds3, ~sum(results(.x, name = "HYPERTENSION_1_vs_0")$padj < 0.05, na.rm = TRUE)) %>%
-    map_df(~as.data.frame(.x) %>% dplyr::rename(qval = .x), .id = "included") %>%
+deseqresults(dds3, names.dset, hypertensionvars = names(dds3)) %>%
+    select(Name, Feature, log2FoldChange) %>%
+    spread(Name, log2FoldChange) %>%
     kable
@@
 
-Number of significant associations `r sum(pull(dset.deseq.signf.na, qval) < 0.05)` for 
-`r dset.deseq.signf.na %>% nrow` features. Number of
-`r dset.deseq.signf.na %>% filter(log2FoldChange > 0) %>% nrow`
-association were positive.
+Significant associations found between blood pressure indices and N:o `r nrow(dset.deseq.signf.na)`
+genera and N:o `r dset.deseq.signf.na %>% pull(Feature) %>% unique %>% length` features. N:o 
+`r dset.deseq.signf.na %>% filter(log2FoldChange > 0) %>% nrow` positive associations found.
 
-```{r deseq dual plot}
+```{r deseq dual plot, echo = FALSE}
 g.deseqgenus <- gtable_frame(ggplotGrob(g.deseq + theme(legend.position = "none")),
@@
                     saltsubset = TRUE)
+pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
 ```
 
-Lactobacillus prevalence in subsample
-
-```{r lactobacillus prevalence}
-pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
-pseq_prevalence(pseq.genus.salt)
-```
+Lactobacillus prevalence in subsample `r sprintf("%.1f%%", 100*pseq_prevalence(pseq.genus.salt))`.
 
@@
 
-```{r scatter plot from deseq}
+```{r scatter plot from deseq, echo = FALSE}
 g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
@@
 
-core.names.flextable %>% save_as_html("cache/s1.html")
+core.names.flextable
 ```
@@
 ```{r supplement table deseq}
-typology.tbls4 <- data.frame(
-    col_keys = colnames(deseq.table),
-    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
-             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
-    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
-    stringsAsFactors = FALSE)
-
 deseq.table <- dset.deseq.signf %>%
@@
 
-typologyformatter(data = deseq.table, font = 9, typology = typology.tbls4) %>%
+typology.deseq <- data.frame(
+    col_keys = colnames(deseq.table),
+    what = c("", rep("Systolic BP", 2), rep("Diastolic BP", 2),
+             rep("Pulse pressure", 2), rep("Mean arterial pressure", 2), rep("Hypertension", 2)),
+    measure = c("", rep(c("Log2FC±SE", "p"), 5)),
+    stringsAsFactors = FALSE)
+
+typologyformatter(data = deseq.table, font = 9, typology = typology.deseq) %>%
     flextable::width(j = 1, width = 1.2) %>%
@@
     flextable::width(j = seq(3, 11, 2), width = 0.55)
-#  %>%   save_as_html("cache/genus.html")
 ```
@@
 
-typologyformatter(data = deseq.table.na, font = 9, typology = typology.tbls4) %>%
+typologyformatter(data = deseq.table.na, font = 9, typology = typology.deseq) %>%
     flextable::width(j = 1, width = 1.2) %>%
@@
     flextable::width(j = seq(3, 11, 2), width = 0.55) 
-#  %>%   save_as_html("cache/species.html")
 ```

### 2020-02-01 articletwo-rrbiome.R

@@ -336,10 +336,13 @@ calculate.betadiversity <- function(pseq, matrix, vars, npermutations = 999) {
                         npermutations = npermutations))
 }
 
-deseqresults <- function(modellist, names.dset, p = 0.05) {
+deseqresults <- function(modellist,
+                         names.dset,
+                         p = 0.05,
+                         hypertensionvars = c("HYPERTENSION")) {
     vars <- c2l(names(modellist))
     lapply(c2l(vars), function(x, models = modellist) {
-        name <- ifelse(x %in% c("HYPERTENSION"), sprintf("%s_1_vs_0", x), x)
+        name <- ifelse(x %in% hypertensionvars, sprintf("HYPERTENSION_1_vs_0", x), x)
         results(models[[x]], name = name) %>%
             as.data.frame %>%
             tibble::rownames_to_column("Feature") }) %>%    

### 2020-02-06 articletwo.Rmd
Wrong panel order

@@
 
+Listing Lactobacillus spp
+
+```{r write supplement}
+lactobacilli.names.incols <- pseq.species %>%
+    taxa %>%
+    mygrep(word = "^Lactobacillus.*Bacteria") %>%
+    renametaxa %>%
+    sort %>%
+    list.partition(parts = 4) %>%
+    mycbind %>%
+    as.data.frame
+
+lactobacilli.names.flextable <- flextable(lactobacilli.names.incols) %>%
+    flextable::fontsize(size = 10, part = "header") %>%
+    flextable::fontsize(size = 10, part = "body") %>%
+    flextable::width(j=1:3, width = 1.8) %>%
+    flextable::align(align = "left", part = "header") %>%
+    flextable::align(align = "left")
+
+lactobacilli.names.flextable 
+```
+
 Listing results for alpha diversity

### 2020-02-06 articletwo-ggplot.R

@@ -62,7 +62,7 @@ plot.beta <- function(diversity,
                       ymax = 0.0010) {
     diversity <- diversity %>% mutate(Qstar = ifelse(beta.p < 0.05, "*", "-"))
     plot.beta.diversity <- ggplot(diversity,
-                                  aes(x = reorder(Name, deseqplotorder(Name)), y = beta.R2)) +
+                                  aes(x = reorder(Name, -deseqplotorder(Name)), y = beta.R2)) +
         geom_bar(stat="identity", color = "black", fill = "gray96") +
         geom_point(aes(y = beta.R2 - 0.00007, shape=Qstar), show.legend=FALSE, color='black', size=14) +
         scale_shape_manual(values=c('*'='*', '-'='')) +

### 2020-03-12 articletwo.Rmd
Lactobacillus excretition instead of concentration

@@
 library(hexbin)
+library(readr)
 ```
@@
 sourcefiles <- c("articletwo-officer.R",
+                 "articletwo-import.R",
                  "articletwo-rrbiome.R",
@@
                                                matrix = bray.dist.m.species,
+                                               npermutations = 999,
                                                vars = list("max" = var.CL,
@@
 ```{r lacto deseq calculation}
-dds.salt <- myDESeq(pseq = pseq.genus,
-                    coretaxa = pseq.genus.coretaxa,
-                    vars = c("NA.", var.CL),
-                    saltsubset = TRUE)
 pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
@@
 Lactobacillus prevalence in subsample `r sprintf("%.1f%%", 100*pseq_prevalence(pseq.genus.salt))`. 
+Levels of dU-Na `r sprintf("%.1f (%.1f)", mean(meta(pseq.genus)$dUNA, na.rm = TRUE), 
+	sd(meta(pseq.genus)$dUNA, na.rm = TRUE))`
 
 ```{r significant results for salt subset}
+dds.salt <- myDESeq(pseq = pseq.genus,
+                    coretaxa = pseq.genus.coretaxa,
+                    vars = c("dUNA", var.CL),
+                    saltsubset = TRUE)
+
 dds.salt %>%
-    results(., name = "NA.", tidy = TRUE) %>%
-    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
+    results(., name = "dUNA", tidy = TRUE) %>%
+    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
+                  lfc_se = sprintf("%.4f±%.4f", log2FoldChange, lfcSE)) %>%
+    select(row, lfc_se, qval) %>%
     filter(row == "Lactobacillus (Bacteria)") %>%
@@
 
-```{r scatter plot from deseq, echo = FALSE}
-g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
-ggsave(file = "cache/gsalt.png", plot = g.salt, height = 5, width = 5, dpi = 300)
+```{r salt subset char}
+df.genus.salt <- saltdset(pseq.genus.salt, dds.salt)
 ```
 
-<img src="cache/gsalt.png" />
+```{r genus salt levels}
+df.genus.salt %>%
+    group_by(group) %>%
+    summarize(mean = round(mean(dUNA), 1),
+              sd = round(sd(dUNA), 1)) %>%
+    kable
+```
+
+Mean urinary salt `r df.genus.salt$dUNA %>% mean` and sd `r df.genus.salt$dUNA %>% sd`.
 
@@
 dds.salt.species <- myDESeq(pseq = pseq.species,
-                        vars = c("NA.", var.CL),
+                        vars = c("dUNA", var.CL),
                         coreterm = "^Lactobacillus.*Bacteria",
@@
 dds.salt.species %>%
-    results(., name = "NA.", tidy = TRUE) %>%
+    results(., name = "dUNA", tidy = TRUE) %>%
     dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
@@
     filter(qval < 0.05) %>%
+    mutate(qval_exp = sprintf("%.4e", qval)) %>%
+    kable
+```
+
+Prevalence for L. paracasei is
+`r pseq_prevalence_format(pseq.species, "Lactobacillus_paracasei (Bacteria)")`.
+and for L. salivarius is 
+`r pseq_prevalence_format(pseq.species, "Lactobacillus_salivarius (Bacteria)")`.
+
+```{r species plot}
+df.species.salt <- deseq.abundances(dds.salt.species) %>%
+    full_join(deseq.abundances(dds.salt), by = "sampleid") %>%
+    left_join(meta(pseq.species) %>% tibble::rownames_to_column("sampleid"), by = "sampleid") %>%
+    mutate_at(vars(contains('Lactobacillus')), .funs = list(group = ~factor(dplyr::ntile(., 4)))) %>%
+    select(dUNA,
+           Lactobacillus_paracasei.Bacteria_group,
+           Lactobacillus_salivarius.Bacteria_group,
+           Lactobacillus.Bacteria_group)
+
+df.species.salt %>%
+    saltboxplot %>%
+    gather(taxa, abundance, -dUNA) %>%
+    ggsave(file = "cache/boxplot-lacto-salt.png", plot = ., height = 3, width = 3.5, dpi = 300, unit = "in")
+```
+
+<img src = "cache/boxplot-lacto-salt.png" />
+
+Urinary sodium quartiles for L. paracei
+
+```{r species salt levels paracasei, echo = FALSE}
+df.species.salt %>%
+    group_by(Lactobacillus_paracasei.Bacteria_group) %>%
+    summarize(mean = round(mean(dUNA), 1),
+              sd = round(sd(dUNA), 1)) %>%
+    kable
+```
+
+Urinary sodium quartiles for L. salivarius
+
+```{r salt levels salivarius, echo = FALSE}
+df.species.salt %>%
+    group_by(Lactobacillus_salivarius.Bacteria_group) %>%
+    summarize(mean = round(mean(dUNA), 1),
+              sd = round(sd(dUNA), 1)) %>%
     kable
@@
 
-```{r write supplement}
+```{r write supplement core names}
 core.names.incols <- pseq.genus.coretaxa %>%
@@
 
-```{r write supplement}
+```{r write supplement lactobacilli}
 lactobacilli.names.incols <- pseq.species %>%

### 2020-03-12 articletwo-ggplot.R

@@ -59,18 +59,21 @@ plot.alpha <- function(diversity,
 
 plot.beta <- function(diversity,
                       fig.textsize = 15,
-                      ymax = 0.0010) {
+                      ymax = 0.0010,
+                      yby = 0.0005,
+                      shift = 0.00007,
+                      ylab = "Coefficient of determination for\nbeta diversity") {
     diversity <- diversity %>% mutate(Qstar = ifelse(beta.p < 0.05, "*", "-"))
     plot.beta.diversity <- ggplot(diversity,
                                   aes(x = reorder(Name, -deseqplotorder(Name)), y = beta.R2)) +
         geom_bar(stat="identity", color = "black", fill = "gray96") +
-        geom_point(aes(y = beta.R2 - 0.00007, shape=Qstar), show.legend=FALSE, color='black', size=14) +
+        geom_point(aes(y = beta.R2 - shift, shape=Qstar), show.legend=FALSE, color='black', size=14) +
         scale_shape_manual(values=c('*'='*', '-'='')) +
         coord_flip() +
         theme_classic(fig.textsize) +
-        ylab("Coefficient of determination for\nbeta diversity") +
+        ylab(ylab) +
         scale_y_continuous(limits = c(0, ymax),
-                           breaks = seq(0, ymax, 0.0005),
+                           breaks = seq(0, ymax, yby),
                            labels = function(x) sprintf("%.2f%%", x*100)) +
         theme(legend.position="right",
               legend.direction="vertical",
@@ -214,7 +217,7 @@ deseqplotorder <- function(x) {
 }
 
 saltplot <- function(dset, ymax = 2) {
-    ggplot(data = dset, aes(x = NA., y = Lactobacillus.Bacteria)) +
+    ggplot(data = dset, aes(x = dUNA, y = Lactobacillus.Bacteria)) +
         geom_point(aes(color = SEX), size = 0.05, position = "jitter") +
         geom_smooth(aes(group = SEX, color = SEX), method='lm', formula= y~x, colour = "black", size = 0.4) +
         scale_x_continuous(limits=c(20, 205), expand = c(0, 0)) +
@@ -239,7 +242,7 @@ myoutlier <- function(list) {
         (list > quantile(list, 0.40) & list < quantile(list, 0.60)) # buggy jitterdodge
 }
 
-saltboxplot <- function(pseq, dds) {
+saltdset <- function(pseq, dds) {
     df.mu <- assays(dds)[["mu"]] %>%
[…]

### 2020-03-12 articletwo-rrbiome.R

@@ -33,56 +33,7 @@ myscale <- function(x){
   (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
 }
 
-filter.phenotype.data <- function(pseq,
-                                  included = c("MEN",
-                                               "BL_AGE",
-                                               "SYSTM",
-                                               "DIASM",
-                                               "BMI",
-                                               "CURR_SMOKE",
-                                               "PREVAL_DIAB",
-                                               "Q57X",
-                                               "BL_USE_RX_C09",
-                                               "BL_USE_RX_C03",
-                                               "BL_USE_RX_C07",
-                                               "BL_USE_RX_C08",
-                                               "NA.",
-                                               "BP_TREAT",
-                                               "PREVAL_CHD",
-                                               "PREVAL_CR_ANYCANC"),
-                                  allowna = c("NA.",
-                                              "BP_TREAT",
-                                              "PREVAL_CHD",
-                                              "PREVAL_CR_ANYCANC")) {
-    meta(pseq) %>%
-        tibble::rownames_to_column(var = "rowname") %>%
-        dplyr::select(rowname, included) %>%
-        tidyr::drop_na(myin(included, allowna, complement = TRUE)) %>%
-        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
-                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
-                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
-                      PULSEPRESSURE = SYSTM - DIASM,
-                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
-                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
-                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
-        dplyr::mutate(oSYSTM = SYSTM,
-                      oDIASM = DIASM,
-                      oPULSEPRESSURE = PULSEPRESSURE,
[…]

### 2020-06-24 articletwo-sensitivityanalysis.Rmd
Final codes

@@
 
+```{r load old, eval = FALSE}
+load("session/session-sensitivity.Rdata")
+```
+
 ### RR biome functions
@@
 
+Self reported BP medication
+
+```{r bp medication use}
+pseq.species %>% meta %>% pull(BP_TREAT) %>% as.factor %>% summary %>% kable
+```
+
 ```{r sensitivity analysis bp treat}
@@
 <img src="cache/sensitivity-alphabeta-bptreat.png"/>
+
+Excluding participants with diabtes, CAD, and any cancer
+
+```{r only healthy}
+pseq.species %>% meta %>%
+    filter(PREVAL_DIAB == 1 | PREVAL_CHD == 1 | PREVAL_CR_ANYCANC == 1) %>%
+    summarize(diab = sum(PREVAL_DIAB == 1),
+              chd = sum(PREVAL_CHD == 1),
+              canc = sum(PREVAL_CR_ANYCANC == 1),
+              any = n()) %>%
+    kable
+```
 

### 2020-06-24 articletwo.Rmd
Final codes

@@
 library(readr)
+library(KEGGREST)
 ```
@@
 
-# Sources
+# Import helper functions
 
@@
 pseq.genus.coretaxa <- coretaxa(pseq.genus, detection = 0.1/100, prevalence = 1/100)
+pseq.genus.core.deseq <-  prune_taxa(pseq.genus.coretaxa, pseq.genus)
 ```
@@
 
-### Bray curtis distance matrix
+## Bray curtis distance matrix
 
@@
 
-### Beta diversity
+## Beta diversity
 
@@
 
-### PCoA
+## PCoA
 
@@
 
-### DeSeq2
+## DeSeq2
+
+Associations for common genera
 
 ```{r deseq2}
-if (!file.exists("rds/dds.rds") || !file.exists("rds/dds3.rds") || !file.exists("rds/ddsNa.rds")) {
-    pseq.genus.core.deseq <-  prune_taxa(pseq.genus.coretaxa, pseq.genus)
-
 if (!file.exists("rds/dds.rds")) {
@@
     saveRDS(dds, file = "rds/dds.rds")
+} else {
+    dds <- readRDS(file = "rds/dds.rds")
 }
+```
+
+The affect of third covariate for hypertension
+
+```{r deseq2 third covariate}    
 if (!file.exists("rds/dds3.rds")) {
@@
     saveRDS(dds3, file = "rds/dds3.rds")
+}  else {
+    dds3 <- readRDS(file = "rds/dds3.rds")
 }
-    if (!file.exists("rds/ddsNa.rds")) {
-        ddsNa <- mclapply(c2l(var.BP), function(x) {
+```
+
+Associations for Lactobacili
+
+```{r deseq2 lactobacilli}
+if (!file.exists("rds/ddsLactobacilli.rds")) {
+    ddsLactobacilli <- mclapply(c2l(var.BP), function(x) {
         myDESeq(pseq = pseq.species,
@@
     }, mc.cores = length(var.BP))
-        saveRDS(ddsNa, file = "rds/ddsNa.rds")        
-    }
+    saveRDS(ddsLactobacilli, file = "rds/ddsLactobacilli.rds")
 } else {
-    dds <- readRDS(file = "rds/dds.rds")
-    dds3 <- readRDS(file = "rds/dds3.rds")
-    ddsNa <- readRDS(file = "rds/ddsNa.rds")
+    ddsLactobacilli <- readRDS(file = "rds/ddsLactobacilli.rds")
 }
@@
 
-# Results
+Subgroup analyses for Lactobacilli
 
-## Characteristics
+```{r heavy subgroup analyses for lactobasilli-BP}
+if (!file.exists("rds/ddsSubLactobacilli.rds")) {
+    ddsSubLactobacilli <- lapply(myDESeq.subsets(), function(subset) {
+        mclapply(c2l(var.BP), function(bp) {
+            myDESeq(pseq = pseq.species,
+                    vars = bp %union% var.CL %difference% subset$exclude,
+                    coreterm = "^Lactobacillus.*Bacteria",
+                    FUN = subset$fun)
+        }, mc.cores = length(var.BP))
+    })
+    saveRDS(ddsSubLactobacilli, file = "rds/ddsSubLactobacilli.rds")
+} else {
+    ddsSubLactobacilli <- readRDS(file = "rds/ddsSubLactobacilli.rds")
+}
+
+#ddsSubLactobacilli <-  readRDS(file = "rds/ddsSub-temp.rds") %>%
+#    purrr::list_modify(drug = readRDS(file = "rds/ddsSubLactobacilli.rds")$drug)
+```
+
+# Characteristics
 
@@
 
-## Microbial diversity
+<details><summary>Chi-square tests and t-tests</summary>
+
+```{r chi squared}
+lapply(c2l(var.CL, var.BP), function(x) {
+    meta(pseq.species) %>%
+        mutate(nagroup = ifelse(is.na(dUNA), 1, 0)) %$%
+        { if (x %in% c("BL_AGE", "BMI", "MAP", "SYSTM", "DIASM", "PULSEPRESSURE"))
+              a <<- t.test(as.formula(sprintf("%s ~ nagroup", x))) 
+          else
+              table(nagroup, .[[x]]) %>%
+                  chisq.test() } %>%
+        tidy}) %>%
+    map_df(~.x, .id = "term") %>%
+    mutate (p.value = pub.p(p.value)) %>%
+    select(term, p.value)
+```
+
+```{r chi squared multilevel}
+lapply(c2l(1, 2, 3), function(x) 
+    meta(pseq.species) %>%
+    mutate(nagroup = ifelse(is.na(dUNA), 1, 0),
+           Q57X = ifelse(Q57X == x, 1, 0)) %$%
+    table(nagroup, Q57X) %>%
+    chisq.test() %>%
+    tidy) %>%
+    map_df(~.x, .id = "term") %>%
+    mutate (p.value = pub.p(p.value)) %>%
+    select(term, p.value)
+```
+
+</details>
+
+# Alpha and beta diversity plots
 
@@
 
-## Principal coordinate analysis
+# Principal coordinate analysis
 
@@
 
-## Associations for genera
+# Associations between genera and blood pressure indices
 
@@
 
-### Associations for hypertension looping covariates
+## N:o associations for hypertension per included 3rd covariate
 
@@
 
-## Associations for Lactobacillus species
+# Associations between Lactobacilli and blood pressure indices
 
 ```{r lactobacillus species results}
-dset.deseq.signf.na <- deseqresults(ddsNa, names.dset)
+dset.deseq.signf.na <- deseqresults(ddsLactobacilli, names.dset)
 g.deseq.na <- deseqheatmap(dset.deseq.signf.na)
@@
 
-## Salt subset
-
-## Association between genera and dU-NA
-
-```{r lacto deseq calculation}
-pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
-```
-
-Lactobacillus prevalence in subsample `r sprintf("%.1f%%", 100*pseq_prevalence(pseq.genus.salt))`. 
-Levels of dU-Na `r sprintf("%.1f (%.1f)", mean(meta(pseq.genus)$dUNA, na.rm = TRUE), 
-	sd(meta(pseq.genus)$dUNA, na.rm = TRUE))`
+# Association between Lactobacillus genus and urinary sodium
 
 ```{r significant results for salt subset}
-dds.salt <- myDESeq(pseq = pseq.genus,
+(dds.salt <- myDESeq(pseq = pseq.genus,
                     coretaxa = pseq.genus.coretaxa,
                     vars = c("dUNA", var.CL),
-                    saltsubset = TRUE)
-
-dds.salt %>%
+                    saltsubset = TRUE)) %>%
     results(., name = "dUNA", tidy = TRUE) %>%
@@
 ```{r salt subset char}
+pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
 df.genus.salt <- saltdset(pseq.genus.salt, dds.salt)
@@
 
-## Association between Lactobacilli and dU-NA
+# Association between Lactobacilli and urinary sodium
+
+Analysis between *Lactobacilli* and urinary salt.
 
 ```{r species salt}
-dds.salt.species <- myDESeq(pseq = pseq.species,
+( dds.salt.species <- myDESeq(pseq = pseq.species,
                         vars = c("dUNA", var.CL),
                         coreterm = "^Lactobacillus.*Bacteria",
-                        saltsubset = TRUE)
-```
-
-Significant results after FDR correction
-
-```{r species salt results}
-dds.salt.species %>%
+                        saltsubset = TRUE) ) %>%
     results(., name = "dUNA", tidy = TRUE) %>%
@@
 
-Prevalence for L. paracasei is
-`r pseq_prevalence_format(pseq.species, "Lactobacillus_paracasei (Bacteria)")`.
-and for L. salivarius is 
-`r pseq_prevalence_format(pseq.species, "Lactobacillus_salivarius (Bacteria)")`.
+```{r test}
+dds.salt.species %>%
+      results(., name = "dUNA", tidy = TRUE) %>%
+    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
+                  row = renametaxa(row)) %>%
+                                        #    filter(qval < 0.05) %>%
+    arrange(qval) %>%
+    mutate(qval_exp = sprintf("%.4e", qval)) %>%
+    kable
+```
+
+```{r species salt}
+dds.salt.species %>%
+    results(., name = "SEX_1_vs_0", tidy = TRUE) %>%
+    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
+                  row = renametaxa(row)) %>%
+    filter(qval < 0.05) %>%
+    mutate(qval_exp = sprintf("%.4e", qval)) %>%
+    kable
+```
+
+The box plot for *Lactobacilli* groups and urinary salt to demonstrate the association.
 
 ```{r species plot}
-df.species.salt <- deseq.abundances(dds.salt.species) %>%
+( df.species.salt <- deseq.abundances(dds.salt.species) %>%
     full_join(deseq.abundances(dds.salt), by = "sampleid") %>%
@@
            Lactobacillus_salivarius.Bacteria_group,
-           Lactobacillus.Bacteria_group)
-
-df.species.salt %>%
+           Lactobacillus.Bacteria_group) ) %>%
+    gather(taxa, abundance, -dUNA) %>%
     saltboxplot %>%
-    gather(taxa, abundance, -dUNA) %>%
     ggsave(file = "cache/boxplot-lacto-salt.png", plot = ., height = 3, width = 3.5, dpi = 300, unit = "in")
@@
 
-Urinary sodium quartiles for L. paracei
+Mean and SD for urinary sodium quartiles
 
 ```{r species salt levels paracasei, echo = FALSE}
-df.species.salt %>%
-    group_by(Lactobacillus_paracasei.Bacteria_group) %>%
-    summarize(mean = round(mean(dUNA), 1),
-              sd = round(sd(dUNA), 1)) %>%
+lapply(c2l(df.species.salt %>% select(contains("Lactobacillus")) %>% colnames),
+       function(x) df.species.salt %>%
+                   dplyr::rename(group = !!x) %>%
+                   group_by(group) %>%
+                   summarize(mean = round(mean(dUNA), 2),
+                             sd = round(sd(dUNA), 2)) ) %>%
+    map_df(~as.data.frame(.x), .id = "taxa")  %>%
+    mutate(taxa = renametaxa(taxa)) %>%
     kable
@@
 
-Urinary sodium quartiles for L. salivarius
+# Subgroup analyses
 
-```{r salt levels salivarius, echo = FALSE}
-df.species.salt %>%
-    group_by(Lactobacillus_salivarius.Bacteria_group) %>%
-    summarize(mean = round(mean(dUNA), 1),
-              sd = round(sd(dUNA), 1)) %>%
+## Sex
+
+```{r deseq dual plot, echo = FALSE}
+g.ddssub.male <- ddsSubLactobacilli$male %>%
+    deseqresults(., names.dset) %>%
+    deseqheatmap(breaks = seq(-2, 2)) %>%
+    { . + theme(legend.position = "none",
+                plot.title = element_text(size = 12, face = "plain")) } %>%
+    myggtitle(., title= "A. Men")
+
+g.ddssub.female <- ddsSubLactobacilli$female %>%
+    deseqresults(., names.dset) %>%
+    deseqheatmap(breaks = seq(-2, 2)) %>%
+    { g.ddssub.female.raw <<- . } %>%
+    { . + theme(legend.position = "none",
+                 plot.title = element_text(size = 12, face = "plain")) } %>%
+    myggtitle(., title= "B. Women")
+
+g.ddssub.legend <- get_legend(g.ddssub.female.raw) %>% as_ggplot
+
+g.deseq.subsample.sex <- arrangeGrob(g.ddssub.male,
+                                     g.ddssub.female,
+                                     g.ddssub.legend ,
+                                     ncol = 3,
+                                     widths = c(2, 2, 1))
+ggsave(file = "cache/deseq-subsample-sex.png",
+       plot = g.deseq.subsample.sex,
+       width = 17,
+       height = 11,
+       dpi = 300,
+       unit = "cm")
+```
+
+## Antihypertensive
+
+```{r deseq dual plot, echo = FALSE}
+g.ddssub.nodrug <- ddsSubLactobacilli$nodrug %>%
+    purrr::list_modify(HYPERTENSION = NULL) %>%
+    deseqresults(., names.dset) %>%
+    deseqheatmap(breaks = seq(-2, 2)) %>%
+    { . + theme(legend.position = "none",
+                 plot.title = element_text(size = 12, face = "plain")) } %>%
+    myggtitle(., title= "A. No antihypertensive medication use")
+
+g.ddssub.drug <- ddsSubLactobacilli$drug %>%
+    purrr::list_modify(HYPERTENSION = NULL) %>%
+    deseqresults(., names.dset) %>%
+    deseqheatmap(breaks = seq(-2, 2)) %>%
+    { . + theme(legend.position = "none",
+                plot.title = element_text(size = 12, face = "plain")) } %>%
+    myggtitle(., title= "B. Antihypertensive medication use")
+
+g.deseq.subsample.drug <- arrangeGrob(g.ddssub.nodrug,
+                                      g.ddssub.drug,
+                                      g.ddssub.legend,
+                                     ncol = 3,
+                                     widths = c(2, 2, 1))
+ggsave(file = "cache/deseq-subsample-drug.png",
+       plot = g.deseq.subsample.drug,
+       width = 19,
+       height = 14.2,
+       dpi = 300,
+       unit = "cm")
+```
+
+# Sodium subsample
+
+```{r subset salt sex}
+( ddsSubLactobacilliSodium <- mclapply(myDESeq.subsets(), mc.cores = 4, function(subset)
+    myDESeq(pseq = pseq.species,
+            vars = "dUNA" %union% var.CL %difference% subset$exclude,
+            coreterm = "^Lactobacillus.*Bacteria",
+            saltsubset = TRUE,
+            FUN = subset$fun)) ) %>%
+    myDESeq.tidy() %>%
+    map_df(~.x, .id = "subgroup") %>%
     kable
+```
+
+```{r subset salt sex}
+ddsSubLactobacilliSodium %>%
+    myDESeq.tidy()) %>%
+    map_df(~.x, .id = "subgroup") %>%
+    kable
+```
+
+## Lactobacilli prevalences
+
+Prevalence
+
+```{r prevalence 0.01%}
+lapply(pseq.species %>% taxa %>% mygrep(word = "^Lactobacillus_", .) %>% c2l, function(x) 
+    pseq_prevalence(pseq.species, taxa = x, limit = 0.01/100)) %>%
+    tibble::enframe() %>% 
+    unnest(cols = value) %>%
+    arrange(desc(value)) %>%
+    kable
+
+females <- pseq.species %>%
+    meta %>%
+    tibble::rownames_to_column("sampleid") %>%
+    filter(SEX == 1) %>%
+    pull(sampleid)
+
+males <- pseq.species %>%
+    meta %>%
+    tibble::rownames_to_column("sampleid") %>%
+    filter(SEX == 0) %>%
+    pull(sampleid)
+
+taxa(pseq.species) %>% mygrep(word = "Salivarius")
+
+pseq_prevalence(prune_samples(males, pseq.species), limit = 0.01/100, taxa = "Lactobacillus_paracasei (Bacteria)")
+pseq_prevalence(prune_samples(females, pseq.species), limit = 0.01/100, taxa = "Lactobacillus_paracasei (Bacteria)")
+
+```
+
+```{r prevalence 0.1%}
+lapply(c("Lactobacillus_paracasei (Bacteria)", "Lactobacillus_salivarius (Bacteria)") %>% c2l, function(x) 
+    pseq_prevalence(pseq.species, taxa = x, limit = 0.1/100)) %>%
+    tibble::enframe() %>% 
+    unnest(cols = value) %>%
+    arrange(desc(value)) %>%
+    kable
+```
+
+# Functional analysis
+
+```{r import KEGG data}
+dset.ko  <- readRDS("data/function_activity_copied_20200213.rds")$ko %>%
+    as.data.frame %>%
+    rownames_to_column(var = "term") %>%
+    gather(sampleid, value, -term) %>% 
+    spread(term, value) %>%
+    mutate_at(vars(starts_with("K")), ~scale(log10(. + 1))) %>%
+    left_join(meta(pseq.species) %>% rownames_to_column(var = "sampleid"), by = "sampleid")
+```
+
+```{r model KEGG}
+ret.ko <- loop.lm(dset = dset.ko,
+               response = "SYSTM",
+               loops = dset.ko %>% colnames %>% mygrep(word = "^K[0-9]+"),
+               covariates = var.CL) %>%
+    loop.results()
+```
+
+```{r process KEGG results}
+ret.ko %>%
+    filter(qval < 0.05) %>%
+    arrange(qval) %>%
+    mutate(mean_ci = sprintf("%.2f (%.2f to %.2f)", estimate, conf.low, conf.high),
+           qval = pub.p(qval)) %>%
+    mutate(desc = mykeggget(term)) %>%
+    select(term, desc, mean_ci, qval) %>%
+    kable
+```
+
+<details><summary>Diagnostic plots</summary>
+
+```{r KEGG qq plots}
+diagnosticqqplot(dset.ko, ret.ko %>% filter(qval < 0.05) %>% pull(term)) %>%
+    ggsave(file = "cache/kegg.png", plot = ., height = 40, width = 10, dpi = 300, unit = "in")
+```
+
+<img src = "cache/kegg.png" />
+
+```{r KEGG scatter plots}
+diagnosticscatterplot(dset.ko,
+                      ret.ko %>% filter(qval < 0.05) %>% pull(term),
+                      SYSTM) %>%
+    ggsave(file = "cache/kegg-scatter.png", plot = ., height = 40, width = 10, dpi = 300, unit = "in")
+```
+
+<img src = "cache/kegg-scatter.png" />
+
+</details>
+
+```{r process KEGG results}
+ret.ko %>%
+    filter(qval < 0.05, estimate < 0) %>%
+    mutate(term = paste0("n-", term),
+           size = paste0("v-", 1/qval)) %>%
+    select(term, size) %>%
+    write_tsv(path = "cache/functree_negative.tsv", col_names = FALSE)
 ```

### 2020-06-24 articletwo-ggplot.R

@@ -1,4 +1,6 @@
 plot.alpha <- function(diversity,
+                       palette = "Blues",
+                       palettedirection = -1,
                        alphabreaks = seq(-1, 0, 0.2),
                        alphalabels = c("(-1.0, -0.8]", "(-0.8, -0.6]", "(-0.6, -0.4]", "(-0.4, -0.2]", "(-0.2, -0.0]")) {
     diversity <- diversity %>%
@@ -14,10 +16,10 @@ plot.alpha <- function(diversity,
                                                   fill = estimate_fac)) +
         geom_bar(stat="identity", color="black") +
         coord_flip() +
-        scale_fill_brewer(palette="Blues",
+        scale_fill_brewer(palette=palette,
                           name = 'Regression coefficient \nin linear model \nfor Shannon index',
                           drop=FALSE,
-                          direction = -1) +
+                          direction = palettedirection) +
         theme_classic(20) +
         geom_point(aes(Name, y=0.5, shape=Qstar), show.legend=FALSE, color='black', size=14) +
         scale_shape_manual(name="",
@@ -179,7 +181,7 @@ plot.diversities <- function(diversities) {
     return(g4)
 }
 
-deseqheatmap <- function(df, legend.name = "Log Base Two\nFold Change") {
+deseqheatmap <- function(df, legend.name = "Log Base Two\nFold Change", breaks = c(-1, -0.5, 0.5, 0, 1)) {
     ggplot(df,
            aes(x = reorder(Name, deseqplotorder(Name)),
                y = ordered(Feature, levels=rev(sort(unique(Feature)))),
@@ -188,8 +190,8 @@ deseqheatmap <- function(df, legend.name = "Log Base Two\nFold Change") {
         geom_tile(colour="white") +
         scale_fill_distiller(palette = "RdBu",
                              name = legend.name,
-                             limits=c(-1, 1),
-                             breaks = c(-1, -0.5, 0.5, 0, 1),
+                             limits = c(min(breaks), max(breaks)),
+                             breaks = breaks,
                              na.value="grey30") +
         coord_fixed(ratio=1) +
         xlab("") +
@@ -263,12 +265,17 @@ saltdset <- function(pseq, dds) {
[…]

### 2020-06-24 articletwo-import.R

@@ -1,0 +1,89 @@
+import_filter_data <- function(file, included = c("MEN",
+                                          "BL_AGE",
+                                          "SYSTM",
+                                          "DIASM",
+                                          "BMI",
+                                          "CURR_SMOKE",
+                                          "PREVAL_DIAB",
+                                          "Q57X",
+                                          "BL_USE_RX_C09",
+                                          "BL_USE_RX_C03",
+                                          "BL_USE_RX_C07",
+                                          "BL_USE_RX_C08",
+                                          "NA.",
+                                          "BP_TREAT",
+                                          "PREVAL_CHD",
+                                          "PREVAL_CR_ANYCANC")) {
+    pseq.full <- readRDS(file)
+    pseq.meta <- pseq.full %>%
+        meta %>%
+        as.data.frame %>%
+        phenotype.filter(included = included) %>%
+        phenotype.definitions 
+    phyloseq::sample_data(pseq.full) <- phyloseq::sample_data(pseq.meta)
+    return(pseq.full)
+}
+
+phenotype.extra <- function(extra, fextra = "pheno/2015_60_Salomaa_Jain_dataFR02_FU17_2020-02-19.txt.gz") {
+    read_tsv(fextra, col_types = cols(
+                         .default = col_double(),
+                         CDT = col_character(),
+                         Sample_ID = col_character(),
+                         Barcode = col_character(),
+                         WESID = col_character(),
+                         WGSID = col_character(),
+                         BATCH = col_character(),
+                         FID = col_character(),
+                         CRP = col_character(),
+                         K_TPKS = col_character(),
+                         K_VKS = col_character(),
+                         K_M1 = col_character(),
+                         K_M2 = col_character(),
+                         APOE_BATCH = col_character())) %>%
+        select(one_of(extra)) %>%
+        filter(!is.na(Barcode))
+}
+
+phenotype.filter <- function(df,
+                            included,
+                             extra = c("Barcode", "U_VIRTSA_YHT"),
+                             allowna = c("NA.",
+                                         "U_VIRTSA_YHT",
[…]

### 2020-06-24 articletwo-officer.R

@@ -47,7 +47,7 @@ characteristicsTable <- function(dset, strata) {
                       printToggle = FALSE,
                       digits = 1,
                       pDigits = 3,
-                      contDigits=1)  %>%
+                      contDigits=2)  %>%
         as.data.frame %>%
         tibble::rownames_to_column(var = "rowname") %>%
         format(justify = "left", trim = TRUE) %>%

### 2020-06-24 articletwo-rrbiome.R

@@ -1,12 +1,26 @@
 # Helper functions for articletwo
 
+`%difference%` <- function(a, b) {
+    a[!a %in% b]
+}
+
+`%intersect%` <- function(a, b) {
+    intersect(a, b)
+}
+
+
+`%union%` <- function(a, b) {
+    c(a, b)
+}
+
 replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
 
-renametaxa <- function (names) {
-    suffix <- ifelse(grepl("BacteriaPlasmid", names), "*", "")
-    base <- ifelse(grepl("_", names),
-           gsub('(.).*_(.*) \\(+(.+)\\)', '\\1. \\2', names),
-           gsub('(.*) \\(+(.+)\\)', '\\1', names))
+renametaxa <- function (names, star = "*") {
+    suffix <- ifelse(grepl("BacteriaPlasmid", names), star, "")
+    base <- case_when(grepl("_.*\\.Bacteria", names) ~ gsub('(.).*_(.*).Bacteria.*', '\\1. \\2', names),
+                      grepl("\\.Bacteria", names) ~ gsub('(.*)\\.Bacteria.*', '\\1', names),
+                      grepl("_", names) ~ gsub('(.).*_(.*) \\(+(.+)\\)', '\\1. \\2', names),
+                      TRUE ~  gsub('(.*) \\(+(.+)\\)', '\\1', names))
     paste0(base, suffix)
 }
 
@@ -91,27 +105,7 @@ calculateadonis <- function(dset,
       }, mc.cores = min(maxcores, length(responses)))
 }
 
-prepare.maaslin <- function (pseq,
-                             tsvfile = "merged_phfinrisk_genus.tsv",
-                             conffile = "maaslin_config.config",
-                             includedvars = c("BL_AGE", "BMI", "SEX"),
-                             readpclrows = "") {
-    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq)))
-    sample.ids <- colnames(pheno)
-    abud.rel <- abundances(pseq)
-    rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
-    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
-    data <- file(conffile , open='wt')
-    pcl.contents <- c("Matrix: Metadata",
[…]
